\blue{...Numerical investigation of viscous flow instabilities through heterogeneous porous media using adaptive and unstructured meshes under different mobility scenarios...}

### 2017-07-28 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
3D case updates and rewrite the conclusions

\subsection{3-D Channel Flows: Investigation of Preferential Flow Path and Formation of Fingers} \label{section:results_dinlet}

\blue{(a) Little intro of the case; }\\
The last of the simulations is a $3$D domain that will help in the description of the finger evolution and the idea of the preferential path using different permeability regions, as these are decribed in figure \ref{fig:3DChannel_PermMesh}.

\blue{(b) set up what are the main issues (theory) of the simulations you are showing in this section; }\\
As already mentioned \textit{viscous fingering} is characterized by the formation of a characteristic pattern, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any discontinuities or instabilities in the flow front.

\blue{(d) then describe the numerical experiment w all BC and IC;}\\
Figure \ref{fig:3DChannel_PermMesh} describes the permeability pattern as this is implied in a $3D$ domain. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a velovity $ u=2.5 cm/sec $. In the areas around the domain a no-slip condition is implied (\ie insulation, no exchange of material or Neumann boundary condition for mass transfer) while at the right hand side of the domain is the outlet. The domain is constructed in such a way so as to be able to capture the preferencial path of flow as this can be seen from figures \ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

\blue{(e) then finally describe the results with a clear conclusion based on observation but backed by theory and/or physics!!}
Figure \ref{fig:3DChannel_planes} gives a comprehensive description of the contours of phase $1$ darcy velocity, in the beggining and just before the end of the simulation. 
While the saturation level for phase $1$, remains between the values of $0.25-0.75$ the main difference between the figures \ref{fig:3DChannel_planes} (a) and (b) is the intensity of the flow as this can been seen by the darcy velocity contours.  Initially most of the contours for phase $1$ are populated and achieve their higher values closer to the inlet region. At this point a comparison between the figures \ref{fig:3DChannel_planes}(a)-\ref{fig:3DChannel_sat}(a) and \ref{fig:3DChannel_planes}(b)-\ref{fig:3DChannel_sat}(f) as they both describe the darcy velocity contours and the saturation front for the same timesteps, is necesserary.

\medskip
As the flow progress, in the $2D$ domain a finger formation will start to appear. This is triggerded initially cause of the permeability differences in the domain. In the $3D$ domain though, the same finger formation does not look the same, as it is more a description of the preferential flow path that may include smaller fingers which will break out of the flow. Figures \ref{fig:3DChannel_sat}(a)-(f) give a description of the preferable flow path in various time-steps alongside with the evolution of the saturation front that now becomes more obvious. Based on the boundary conditions that implemented, the preferencial path of the flow can been seen. In a large scale enhanced oil-recovery operation this might have direct impact on the in-situ hydrocarbons located in lower permeability areas.

\medskip
Thus macroscopic sweep efficiency is affected by the heterogeneity in the reservoir which controls the spatial distribution of porosity and permeability. The permeability depends on the number, size and connectivity of the pores. An on purpose implementation of significant high permeability zones (compared with the lower ones) also know as higher permeability channels or layers in the set up of the model, allows our test-cases to give a good approximation of the so called "thief zones", something that is common in heterogeneous geological formations.

\medskip 
The evolution and results of the saturation front and the idea of preferential flow path in the domain is followed by a more detail representation as this can be found in figure \ref{fig:3DChannel_Isosurf}. For the same timesteps of the saturation front evolution an inside look on the domain provides with a more detailed description of not only the flow path itself as well as the minor fingers that may be created.

\medskip
At this point it is proper to review the mechanisms of viscous fingering (\ie shielding, spreading and tip-splitting, \citet{homsy_1987}). In all the previous cases, as surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability. As a result of shielding, one of these fingers will eventually outgrow the other and, will then spread to occupy the appropriate width of the domain ahead of it. Following this process, it reaches a width that is again unstable and the pattern repeats. So far shielding and spreading were dominating the flow but using this this domain and conditions the mechanics of tip-splitting along a front can be understood at least qualitatively.

@@

\begin{enumerate}
\item[1.] After having separate the cross flow pattern from the viscous finger one, the findings show that the displacing fluid tends to flow easier towards the zones with higher values of permeability, as these can been seen from evolution of fingers in the upper layers of the domains or from cross-section (\red{i want to say that in hele-shaw cells this can be seen from above, longitudinal section? or frame plan}) througout the Hele-shaw domains. 
\item[2.] Even though different permeability patterns will initially make the uniform front to collapse, it is the \textit{viscocity ratio} the key parameter and the capillary forces as these are related with surface tension that will eventually lead the flow instabilities. The higher the viscocity ratio the easier it is for the uniform front of the injected fluid to break into fingers that eventually will decrease the recovery of hydrocarbon content.
\item[3.] The increased viscosity ratio increases the computational cost and requires the use of smaller time-step sizes. 
\item[4.] Properties such as porosity and saturation can be readily up-scaled as they can be included in the element boundaries. The domains used in these test-cases have the ability to describe the fingering mechanism and the development of the flow \ie the latter case clearly describes the linear tip-splitting finger behaviour inside the domain.
\end{enumerate}

>>>>

\subsection{3-D channel flows} \label{section:results_3D}

The last of the simulations introduces a $3$D domain that will showcase the idea of the preferential flow path and finger growth using different permeability regions, as these are decribed in figure \ref{fig:3DChannel_PermMesh}. In large scale geological formations preferential flow allows much faster contaminant transport and creates significant consequences for ground-water quality.

\medskip
As already mentioned in section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of a characteristic pattern, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any discontinuities or instabilities in the flow front. 
Figure \ref{fig:3DChannel_PermMesh} describes the permeability pattern as this is implied in a $3D$ domain. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a velovity $ u=2.5 cm/sec $. In the areas around the domain a no-slip condition is implied (\ie insulation, no exchange of material or Neumann boundary condition for mass transfer) while at the right hand side of the domain is the outlet. The domain is constructed in such a way so as to be able to focus and capture the preferential path of flow as this can be seen from figures \ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

\medskip
Figure \ref{fig:3DChannel_planes} gives a comprehensive description of the contours of phase $1$ darcy velocity, during the beggining and just before the end of the simulation. While the saturation level for phase $1$, remains between the values of $0.25-0.75$ the main difference between the figures \ref{fig:3DChannel_planes} (a) and (b) is the intensity of the flow as this can be given by the darcy velocity contours. In the beginning of the simulation most of the contours for phase $1$ are populated and achieve their higher values closer to the inlet region. At this point a comparison between the figures \ref{fig:3DChannel_planes}(a)-\ref{fig:3DChannel_sat}(a) and \ref{fig:3DChannel_planes}(b)-\ref{fig:3DChannel_sat}(f) as they both describe the darcy velocity contours and the saturation front for the same timesteps, is necesserary.

\medskip
In previous sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}, a $2$D domain was used to describe one or more finger formations. These flow instabilities were triggerded because of the permeability differences in the domain. Using the $3D$ domain, the difference between a flow instability and the preferential flow path that may include smaller fingers is more obvious. Figures \ref{fig:3DChannel_sat}(a)-(f) give a description of the preferable flow path in various time-steps alongside with the progression of the saturation front that now becomes more obvious. Based on the boundary conditions that implemented, the preferencial path of the flow can been seen. Heterogeneity in the reservoir controls the spatial distribution of porosity and permeability. The permeability depends on the number, size and connectivity of the pores. An on purpose implementation of significant high permeability zones (compared with the lower ones) also know as higher permeability channels, allows our test-cases to give a good approximation of the so called "thief zones".

\medskip 
The saturation front progress and the idea of preferential flow path is followed by a more detail representation as this can be found in figure \ref{fig:3DChannel_Isosurf}. For the same time-steps an inside view on the domain provides with a more detailed description of not only the flow path itself as well as the minor fingers that may be created. At this point resolution plays an important role on capturing these patterns in detail and is also a limititation that needs to be overcome.

\medskip
A quick review to the mechanisms of viscous fingering (\ie shielding, spreading and tip-splitting, \citet{homsy_1987}) is required. In all the $2$D previous cases (sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}), as surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability. Through \textit{shielding} mechanism, one of these fingers will eventually outgrow the other and, will then spread (\textit{spreading}) to occupy the appropriate width of the domain ahead of it. Following this process, it reaches a width that is again unstable and the pattern repeats. So far shielding and spreading were dominating the flow but using the current $3$D domain and because of the lack of high resolution, the conditions and the mechanics of tip-splitting along a front can be qualitatively understood.

@@

\begin{enumerate}
\item[1.] After having separate the cross flow pattern from the viscous finger one, the findings show that the displacing fluid tends to flow easier towards the zones with higher values of permeability, as these can been seen from evolution of fingers in the upper layers of the domains or from longitudinal section througout the Hele-shaw domains. 
\item[2.] Even though different permeability patterns will initially make the uniform front to collapse, it is the \textit{viscocity ratio} the key parameter and the capillary forces as these are related with surface tension that will eventually lead the flow instabilities. The higher the viscocity ratio the easier it is for the uniform front of the injected fluid to break into fingers that eventually will decrease the recovery of hydrocarbon content.
\item[3.] The increased viscosity ratio increases the computational cost and requires the use of smaller time-step sizes. 
\item[4.] Properties such as porosity and saturation can be readily up-scaled as they can be included in the element boundaries. The domains used in these test-cases have the ability to describe the fingering mechanism and the development of the flow \ie the latter case clearly describes the linear tip-splitting finger behaviour inside the domain.
\end{enumerate}
